\documentclass[11pt,a4paper]{scrartcl}
\typearea{12}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{pstricks}
\usepackage{listings}

\usepackage{tikz}

\usepackage{pgf}
\usepackage[utf8]{inputenc}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}


\tikzset{
    area/.style={
      circle,
           draw=black, very thick,
           inner sep=2pt,
           text centered,
           },
}


\tikzset{
    center/.style={
      circle,
           draw=black, very thick,
           inner sep=2pt,
           text centered,
           minimum size=2.5cm,
           },
}


\tikzset{
    neuron/.style={
      circle,
           draw=red, very thick,
           inner sep=2pt,
           text centered,
           minimum size=2.5cm,
           },
}


\tikzset{
    applic/.style={
      circle,
           draw=green, very thick,
           inner sep=2pt,
           text centered,
           minimum size=2.5cm,
           },
}



\tikzset{
    neuronfaded/.style={
      circle,
           draw=red, thin,
           inner sep=2pt,
           text centered,
           minimum size=2.5cm,
           },
}


\tikzset{
    applicfaded/.style={
      circle,
           draw=green, thin,
           inner sep=2pt,
           text centered,
           minimum size=2.5cm,
           },
}






\lstset{language=python}
\pagestyle{headings}
\markright{Computational Neuroscience - course plan}

\begin{document}
\section*{Course plan}
\begin{center}
\begin{tikzpicture}

\node[center,text height=0.35cm](nt){neurons types};

\node[center,text height=0.35cm,right = 2cm of nt](sc){some maths};
\node[center,text height=0.35cm,above = 2cm of sc](tb){the brain};


\path (tb) edge[->,line width=2pt](sc);
\path (sc) edge[->,line width=2pt](nt);



\node[neuron,text height=0.35cm,above = 2cm of nt](on){MP neurons};
\node[applic,text height=0.35cm,above = 2cm of on](hpc){hippocampus};

\path (nt) edge[->,line width=2pt](on);
\path (on) edge[<->,line width=2pt](hpc);

\node[neuron,text height=0.35cm,above left = 2cm of nt](rn){rate neurons};
\node[applic,text height=0.35cm,above left = 2cm of rn](vis){vision};

\path (nt) edge[-,line width=2pt,style=dashed](rn);
\path (rn) edge[<->,line width=2pt](vis);


\node[neuron,text height=0.35cm,left = 2cm of nt](ln){LIF neurons};
\node[applic,text height=0.35cm,left = 2cm of ln](pla){features};

\path (nt) edge[-,line width=2pt,style=dashed](ln);
\path (ln) edge[<->,line width=2pt](pla);


\node[neuron,text height=0.35cm,below left = 2cm of nt](hn){HH neurons};
\node[applic,text height=0.35cm,below left = 2cm of hn](pat){patterns};

\path (nt) edge[-,line width=2pt,style=dashed](hn);
\path (hn) edge[<->,line width=2pt](pat);


\node[neuronfaded,text height=0.35cm,below = 2cm of nt](nm){\textcolor{gray}{neural mass}};
\node[applicfaded,text height=0.35cm,below = 2cm of nm](eeg){\textcolor{gray}{fmri / eeg}};

\path (nt) edge[-,line width=1pt,draw=gray,style=dashed](nm);
\path (nm) edge[<->,line width=1pt,draw=gray](eeg);

\path (on) edge[->,line width=2pt](rn);
\path (rn) edge[->,line width=2pt](ln);
\path (ln) edge[->,line width=2pt](hn);
\path (hn) edge[->,line width=1pt,draw=gray](nm);


\end{tikzpicture}
\end{center}

\section*{Key to the plan}


\begin{enumerate}[label=(\alph*)]
\item \textbf{the brain:} A quick and easy outline introduction to the brain and neuroscience.
\item \textbf{some math:} An introduction to scientific computing and differential equations.
\item \textbf{neuron types:} An overview of neuronal modelling.
\item \textbf{\textcolor{red}{MP neurons:}} The McCulloch Pitts model of neurons, simple synapses.
\item \textbf{\textcolor{green}{hippocampus:}} Hebbian plasticity, description of the hippocampus and auto-associative memory.
\item \textbf{\textcolor{red}{rate neurons:}} The rate model of neurons, including
  receptive fields.
\item \textbf{\textcolor{green}{vision:}} The visual pathway; V1, receptive fields in V1 and sparse coding.
\item \textbf{\textcolor{red}{LIF neurons:}} Spiking, spike triggered averages and time histograms, the leaky integrate and fire neuron.
\item \textbf{\textcolor{green}{features:}} Spike timing dependent plasticity and feature extraction.
\item \textbf{\textcolor{red}{HH neurons:}} Ion channels and Hodgkin-Huxley neurons;
  Morris-Lecar and other models.
\item \textbf{\textcolor{green}{patterns:}} Some ideas from dynamical systems, central pattern generators.
\item \textbf{\textcolor{red}{neural mass:}} Neural mass models and field equations for neurons.
\item \textbf{\textcolor{green}{fmri / eeg:}} From neurons to the brain; modelling the interactions of brain regions.
\end{enumerate}

The last two sections are aspirational, it is likely there will not be time to discuss them and they will not be examined.

\section*{Rough lecture list}
This is a rough guide, it might change as the course progresses.

\begin{enumerate}
\item Introduction to the course and to the brain. (a)
\item More on the brain. (a)

\item Yet more on the brain. (a)
\item Scientific programming in Python or Julia. (b)

\item Introduction to differential equations. (b)
\item Numerical solutions to differential equations. (b)

\item Modelling neurons. (c)
\item The McCullough-Pitts neuron and Hebbian plasticity. (d)

\item The Hippocampus. (e)
\item CA3 as a auto-associative memory, pattern seperation. (e)

\item Firing rates, dealing with neuronal data, receptive fields. (f)
\item The visual system. (g)

\item V1 and sparse coding. (g)
\item Spikes and analysing spike date. (h)

\item Simple models of neurons: leaky integrate and fire neurons. FI curves (h)
\item Synapses and synaptic plasticity. (i)

\item STDP and feature extraction. (i)
\item Ion channels. (j)

\item The Hodgkin-Huxley equation and spikes. (j)
\item Dynamical systems approachs, the Morris-Lecar model, phase diagrams. (j/k)

\item Central pattern generators, bursting. (k)
\item Spare lecture.

\end{enumerate}

\end{document}

