\documentclass[12pt]{article}
\usepackage{amsfonts, epsfig}

\usepackage{graphicx}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lfoot{\texttt{github.com/conorhoughton/COMS30127}}
\lhead{Computation Neuroscience - Coursework 3 - integrate and fire}
\rhead{\thepage}
\cfoot{}
\begin{document}

\section*{Coursework 3}

Write a brief report, no longer than two pages, probably shorter,
including the figures and the comments specified above; submissions
exceeding the page limit will be rejected, I will take a dim view of
super-narrow margins or tiny fonts. Submit it in the pdf format
together with the Python or Julia code by the deadline. Remember,
provided it is in good time, I am happy to answer questions about
Python and to help debug faulty code. Submit the report as pdf; make
sure to set the size of the graphs in your plotting programme, not by
shrinking the graph to fit in the report, this makes the figure
legends and axis numbers tiny and is super annoying. You don't have to
plot in Python or Julia, it is fine to output a data file and plot
using gnuplot or whatever.

\subsection*{Question 1}

Simulate an integrate and fire model with the following parameters for
1 s: $\tau_m = 10 $ms, $E_L = V_r = -70$ mV, $V_t = -40$ mV, $R_m= 10$
M$\Omega$, $I_e = 3.1 $ nA. Use Euler's method with timestep $\delta t
= 1$ ms. Here $E_L$ is the leak potential, $V_r$ is the reset voltage,
$V_t$ is the threshold, $R_m$ is the membrane resistance, that is one
over the conductance, and $\tau_m$ is the membrane time constant. Plot
the voltage as a function of time. For simplicity assume that the
neuron does not have a refractory period after producing a spike. You
do not need to plot spikes - once membrane potential exceeds
threshold, simply set the membrane potential to $V_r$. [3 marks for
  correct; 1 for some spiking but not at correct rate etc, 1 mark lost
  if graphs hard to read, figure legends missings, units missing]

\subsection*{Question 2}

Simulate two neurons which have synaptic connections between
  each other, that is the first neuron projects to the second, and the
  second neuron projects to the first. Both model neurons should have
  the same parameters: $\tau_m = 20$ ms, $E_L = -70$ mV $V_r = -80$ mV
  $V_t = -54$ mV $R_mI_e = 18$ mV and their synapses should also have
  the same parameters: $R_m \bar{g}_s = 0.15$, $P = 0.5$, $\tau_s= 10$
  ms; don't get confused by being given $R_m\bar{g}_s$ rather than
  $\bar{g}_s$ on its own, to get $\tau_m$ rather than the capacitance
  on the left hand side of the integrate and fire equation everything
  is multiplied by $R_m$. For simplicity take the synaptic conductance
\begin{equation}
g_s=\bar{g}_s s
\end{equation}
to satisfy
\begin{equation}
\tau_s\frac{ds}{dt}=-s
\end{equation}
with a spike arriving causing $s$ to increase by $P$. This is
equivalent to the simple synapse model in the lectures. Simulate two
cases: a) assuming that the synapses are excitatory with $E_s = 0$ mV,
and b) assuming that the synapses are inhibitory with $E_s = -80$
mV. For each simulation set the initial membrane potentials of the
neurons $V$ to different values chosen randomly from between $V_r$ and
$V_t$ and simulate 1 s of activity. For each case plot the voltages of
the two neurons on the same graph (with different colours). [7 marks;
  3 for correct general approach but neurons connected up wrong,
  synapses coupled to the wrong voltage etc, 2 each for the inhibitory
  and excitatory graphs correct, 2 marks lost for problems with the
  graphs, eg no units, tiny script, cut and pasted screen shot, etc].

\subsection*{COMSM2127}

\begin{enumerate}

\item Compute analytically the minimum current $I_e$ required for the
  neuron with the above parameters to produce an action
  potential. [2 marks]

\item Simulate the neuron for 1 s for the input current with amplitude
  $I_e$ which is 0.1 [nA] lower than the minimum current computed
  above, and plot the voltage as a functions of time. [1 mark]

\item Simulate the neuron for 1s for currents ranging from 2 [nA] to 5
  [nA] in steps of 0.1 [nA]. For each amplitude of current count the
  number of spikes produced, that is the firing rate. Plot the firing
  rate as the function of the input current. [2 marks]

\end{enumerate}

\end{document}

